---
title: "Task 3"
author: "Deanna Elliott"
date: "10/23/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(digits = 2)

#attach packages

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(lubridate)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER

```{r}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Visualization ##

Create a species-faceted histogram of urchin size and count with caption

```{r histogram}

ggplot(data = urchins,
       aes(x = size, fill = common_name)) +
           scale_fill_manual(values = c("darkorchid3", "firebrick3"),
                             name = "Species") +
  xlim(0, 15) +
  geom_histogram(bins = 10) +
  facet_wrap(~ common_name) +
  labs(x = "Diameter (cm)",
           y = "Count")
```

**Figure 1. Urchin count by size and species.** This histogram shows observations of the number and diameter in cm of purple and red sea urchins (*Strongylocentrotus purpuratus* and *Mesocentrotus franciscanus*) from 5 sites along the Santa Barbara Coastal LTER.

**

Create a summary table of mean, stdev, median, and sample size for ech urchin species

```{r stats table}

urchin_summary <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size_cm = mean(size, na.rm = TRUE), 
            sd_size_cm = sd(size, na.rm = TRUE),
            median_size_cm = median(size, na.rm = TRUE),
            sample_size = n())

urchin_summary %>% 
  kbl(col.names = c("Species Common Name", "Mean Size (cm)", "Standard Deviation (cm)",
                    "Median Size (cm)", "Sample Size (n)"),
      caption = "<b>Table 1. Summary statistics for urchin sizes.</b> Mean, standard deviation, and median in cm and sample size of purple and red sea  urchins (<i>Strongylocentrotus purpuratus</i> and <i>Mesocentrotus franciscanus</i>) observed at Santa Barbara Coastal LTER.") %>% 
  kable_classic()
```

### Takeaways ###

Based on the histogram and the summary statistics, red urchins are, on average, approximately 2.5 cm--or 35%--larger than purple urchins. Red urchins also have a larger size range compared to purple, as their distribution in the histogram is more widely spread. The central tendencies of each species are expressed approximately equally by the mean and median (4.5 cm and 4.5 cm for purple; 6.8 cm and 7.0 cm for red).

## Data Wrangling ##

Filter to only purple urchins, change "date" column class to date from character, add "year" and "site_full" columns

```{r filter purple}

purple_urchins <- urchins %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))
```

## Data Visualization ##

Group purple urchins by site and year, find sample sizes, plot 

```{r purple table, include = FALSE}

purple_sample <- purple_urchins %>% 
  group_by(site, year) %>% 
  summarize(sample_size = n())

kbl(purple_sample)
```

```{r purple graph}

ggplot(purple_sample,
       aes(x = year, y = sample_size)) +
  geom_col() +
  facet_wrap(~ site) +
  labs(x = "Year",
       y = "Urchin Count")
```

**Figure 2. Urchin count by site and year.** This bar graph shows observations of the sample size of purple sea urchins (*Strongylocentrotus purpuratus*) across several years at 5 sites along the Santa Barbara Coastal LTER.

**

Based on the above graph, it seems that the purple urchin population has gone down in recent years. However, we must take into account the time and effort put into data collection each year; if, for example, one year had fewer volunteers for data collection, there may be fewer observations just by virtue of less manpower. Other possible explanations for population changes include predator levels, ocean chemistry changes, etc.

## End Task ##

