---
title: "Task 3"
author: "Deanna Elliott"
date: "10/23/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

#attach packages

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER

```{r}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Visualization ##

```{r histogram}

ggplot(data = urchins,
       aes(x = size, fill = common_name)) +
           scale_fill_manual(values = c("darkorchid3", "firebrick3"),
                             name = "Species") +
  geom_histogram(binwidth = 10) +
  facet_wrap(~ common_name) +
  labs(x = "Diameter (cm)",
           y = "Count")
```

**Figure 1. Urchin count by size and species.** This histogram shows observations of the number and diameter in cm of purple and red sea urchins (*Strongylocentrotus purpuratus* and *Mesocentrotus franciscanus*) from 5 sites along the Santa Barbara Coastal LTER.

**

**Table 1. Summary statistics for urchin sizes.** Mean, standard deviation, and median in cm and sample size of purple and red sea  urchins (*Strongylocentrotus purpuratus* and *Mesocentrotus franciscanus*) observed at Santa Barbara Coastal LTER.

```{r stats table}

urchin_summary <- urchins %>% 
  group_by(common_name) %>% 
  summarize(mean_size_cm = mean(size, na.rm = TRUE), 
            sd_size_cm = sd(size, na.rm = TRUE),
            median_size_cm = median(size, na.rm = TRUE),
            sample_size = n())

urchin_summary %>% 
  kbl(col.names = c("Species Common Name", "Mean Size (cm)", "Standard Deviation (cm)",
                    "Median Size (cm)", "Sample Size (n)")) %>% 
  kable_classic()
```

TAKEAWAY GOES HERE

```{r filter purple}

```

```{r purple table, include = FALSE}

```

```{r purple graph}

```

GRAPH QUESTIONS

## End Task ##