---
title: "Task 4"
author: "Deanna Elliott"
date: "10/26/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER 

```{r csv}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Wrangling ##

Compare mean size of purple urchins in both treatment groups at Naples Reef in 2010

```{r purple subset}

purple_NAPL <- urchins %>%
  filter(site == "NAPL") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(year == "2010")

purple_summary <- purple_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean_size_cm = mean(size))

purple_summary
```

## Data Exploration ##

```{r hist/qq}

ggplot(data = purple_NAPL,
       aes(x = size)) + 
  geom_histogram(bins = 10) +
  facet_wrap(~ treatment)

purple_NAPL_con <- purple_NAPL %>% 
  filter(treatment == "CONTROL")

purple_NAPL_ann <- purple_NAPL %>% 
  filter(treatment == "ANNUAL")

qqnorm(purple_NAPL_con$size)

qqnorm(purple_NAPL_ann$size)
```

