---
title: "Task 4"
author: "Deanna Elliott"
date: "10/26/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER 

```{r csv}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Wrangling ##

Compare mean size of purple urchins in both treatment groups at Naples Reef in 2010

```{r purple subset}

purple_NAPL <- urchins %>%
  filter(site == "NAPL") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(year == "2010")

purple_summary <- purple_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean_size_cm = mean(size))
```

## Data Exploration ##

Create exploratory histogram and qq plots for both treatments in purple_NAPL

```{r hist}

ggplot(data = purple_NAPL,
       aes(x = size)) + 
  geom_histogram(bins = 10) +
  facet_wrap(~ treatment)
```

Purple urchin size distributions seem more skewed toward the smaller side for both treatments, though the annual kelp removal treatment definitely has a higher volume of small urchins. On the other hand, the control treatment with no kelp removal has a higher volume of larger urchins.

```{r qq control}
purple_NAPL_con <- purple_NAPL %>% 
  filter(treatment == "CONTROL")

qqnorm(purple_NAPL_con$size)
```

The control treatment is not normal, based on the qq plot. As shown in the histogram, the data is negatively skewed.

```{r qq annual}

purple_NAPL_ann <- purple_NAPL %>% 
  filter(treatment == "ANNUAL")

qqnorm(purple_NAPL_ann$size)
```

The annual treatment is also not normal, based on its qq plot. The data is even more negatively skewed, as seen in the histogram.


## Statistical Summary Table ##

Create a finalized table of summary stats for purple_NAPL subset

```{r summary table}

purple_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(size, na.rm = TRUE),
            stdev = sd(size, na.rm = TRUE),
            median = median(size, na.rm = TRUE),
            sample_size = n()) %>% 
  kbl(col.names = c("Treatment", "Mean Size (cm)", "Standard Deviation (cm)", "Median Size (cm)", "Sample Size (n)"),
         caption = "<b>Table 1. Summary statistics for urchin sizes by treatment.</b> Mean, standard deviation, and median in cm with sample sizes of two treatments of purple urchins (<i>Strongylocentrotus purpuratus</i>) observed at Naples Reef in the Santa Barbara Coastal LTER in 2010.") %>% 
  kable_classic()
```

## End Task ##

