---
title: "Task 3"
author: "Deanna Elliott"
date: "10/23/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

#attach packages

library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER

```{r}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Wrangling ##

```{r histogram}

```

```{r stats table}

```

TAKEAWAY GOES HERE

```{r filter purple}

```

## Data Visualization ##

```{r purple table, include = FALSE}

```

```{r purple graph}

```

GRAPH QUESTIONS

## End Task ##