---
title: "Task 4"
author: "Deanna Elliott"
date: "10/26/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
```

## Overview ##

This code wrangles and visualizes data of red and purple urchins sizes observed at 5 sites along the Santa Barbara coast.

**Data Citation:** Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b. 

## Read In Data ##

urchins.csv contains observations about size frequencies of red and purple sea urchins in the Santa Barbara Coastal LTER 

```{r csv}

urchins <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

## Data Wrangling ##

Compare mean size of purple urchins in both treatment groups at Naples Reef in 2010

```{r purple subset}

purple_NAPL <- urchins %>%
  filter(common_name == "Purple Urchin") %>% 
  filter(site == "NAPL") %>% 
  mutate(date = mdy(date)) %>% 
  mutate(year = year(date)) %>% 
  filter(year == "2010")

purple_summary <- purple_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean_size_cm = mean(size))
```

## Data Exploration ##

Create exploratory histogram and qq plots for both treatments in purple_NAPL

```{r hist}

ggplot(data = purple_NAPL,
       aes(x = size)) + 
  geom_histogram() +
  facet_wrap(~ treatment)
```

Purple urchin size seems more skewed toward the smaller side for the annual kelp removal treatment. On the other hand, the control treatment with no kelp removal has a more even distribution, though it is also skewed slightly positively.


```{r qq plots}

ggplot(data = purple_NAPL, aes(sample = size)) +
  geom_qq() + 
  facet_wrap(~ treatment)
```

The annual treatment is not normal, based on its histogram and qq plot. The data is quite positively skewed. On the other hand, I would say the control data is normal, or normal enough, based on the plots. Though it is slightly positively skewed, the data follows the normal curve fairly well in the histogram.


## Statistical Summary Table ##

Create a finalized table of summary stats for purple_NAPL subset

```{r summary table}

purple_NAPL %>% 
  group_by(treatment) %>% 
  summarize(mean = mean(size, na.rm = TRUE),
            stdev = sd(size, na.rm = TRUE),
            median = median(size, na.rm = TRUE),
            sample_size = n()) %>% 
  kbl(col.names = c("Treatment", "Mean Size (cm)", "Standard Deviation (cm)", "Median Size (cm)", "Sample Size (n)"),
         caption = "<b>Table 1. Summary statistics for urchin sizes by treatment.</b> Mean, standard deviation, and median in cm with sample sizes of two treatments of purple urchins (<i>Strongylocentrotus purpuratus</i>) observed at Naples Reef in the Santa Barbara Coastal LTER in 2010.") %>% 
  kable_classic()
```

## End Task ##

